---
title: "Group Project 2"
author: "Erica Fossee & Team"
date: "2/21/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Open data and check variable types, NAs
```{r}
hotel<-read.csv("hotel_bookings.csv", stringsAsFactors = TRUE)
#hotel$reservation_status_date<- as.Date(hotel$reservation_status_date)
```
The following variables can be made into factors:  is_canceled, is_repeated_guest
```{r}
hotel$is_canceled<-as.factor(hotel$is_canceled)
hotel$is_repeated_guest<-as.factor(hotel$is_repeated_guest)
```
The following columns had NAs: children. We will assume they have no kids. 
```{r}
hotel$children<-ifelse(is.na(hotel$children), 0, hotel$children)
```

The following descriptive analyses were done to get a better sense of the data. A new column called "cancelednumeric" was created to convert the prediction variable (is_canceled) to numeric.
```{r}
summary(hotel)
str(hotel)
hotel$cancelednumeric<-as.numeric(hotel$is_canceled)
hotel$cancelednumeric<-ifelse(hotel$cancelednumeric=="1",0,1)

table(hotel$arrival_date_month)
table(hotel$children)
hist(hotel$children)
tapply(hotel$cancelednumeric, hotel$hotel, mean, na.rm = TRUE)
hist(hotel$arrival_date_day_of_month)
tapply(hotel$children, hotel$cancelednumeric, mean, na.rm = TRUE)
tapply(hotel$lead_time, hotel$is_canceled, mean, na.rm = TRUE)
tapply(hotel$cancelednumeric, hotel$adults, mean, na.rm = TRUE)
table(hotel$customer_type)
table(hotel$customer_type, hotel$distribution_channel)
tapply(hotel$cancelednumeric, hotel$customer_type, mean, na.rm = TRUE)
table(hotel$is_canceled, hotel$reservation_status)
table(hotel$cancelednumeric, hotel$previous_cancellations)
table(hotel$is_canceled, hotel$is_repeated_guest)
table(hotel$is_canceled, hotel$deposit_type)
table(hotel$is_canceled, hotel$agent)
table(hotel$arrival_date_year)
tapply(hotel$cancelednumeric, hotel$arrival_date_month, mean, na.rm = TRUE)
```
July and August had the most cancellations, December and January had the least cancellations
More cancellations for city hotels (42%) compared to resort hotels (28%)
There is a drop in hotel arrival date at the end of the month
Bookings made further in advance had more cancellations than bookings made sooner
Parties with more adults were more likely to cancel
Most reservations were made by TA/TO (travel agency)
Groups were least likely to cancel and transient parties were most likely to cancel
Much more cancellations for non-refundable deposits- possibly due to another variable impacting relationship or hotel purposefully giving customers who are likely to cancel a non refundable option
People with more cancellations were more likely to cancel than not cancel
Possibly want to remove the column about customer type or only look at transient parties 
Remove the reservation_status column because this includes the prediction outcome 
Possibly want to predict cancelled, no show, and checked out; or remove no show or remove canceled 


Decrease the number of factor levels for the following variables:country, agent, company, date (?)

agent
Top 10: PRT, GBR, FRA, ESP, DEU, ITA, IRL, BEL, BRA, NLD
```{r}
levels(hotel$country)
levels(hotel$country) <- c(levels(hotel$country),"OTHER")
sort(table(hotel$country), decreasing = T)[1:10]
hotel$country_new<-hotel$country
hotel$country_new[hotel$country!="PRT" & hotel$country!="GBR" & hotel$country!="FRA" & hotel$country!="ESP" & hotel$country!="DEU" & hotel$country!="ITA" & hotel$country!="IRL" & hotel$country!="BEL" & hotel$country!="BRA" & hotel$country!="NLD"]<-"OTHER"
hotel$country<-hotel$country_new
hotel$country<-droplevels(hotel$country)
hotel$country_new<-NULL
```

Agent
```{r}
levels(hotel$agent)
levels(hotel$agent) <- c(levels(hotel$agent),"OTHER")
sort(table(hotel$agent), decreasing = T)[1:10]
hotel$agent_new<-hotel$agent
hotel$agent_new[hotel$agent!="9" & hotel$agent!="NULL" & hotel$agent!="240" & hotel$agent!="1" & hotel$agent!="14" & hotel$agent!="7" & hotel$agent!="6" & hotel$agent!="250" & hotel$agent!="241" & hotel$agent!="28"]<-"OTHER"
table(hotel$agent_new)
hotel$agent <- hotel$agent_new
hotel$agent<-droplevels(hotel$agent)
hotel$agent_new<-NULL
```

Company
NULL 40 223 67 45 153 174 219 281 154
```{r}
levels(hotel$company)
levels(hotel$company) <- c(levels(hotel$company),"OTHER")
sort(table(hotel$company), decreasing = T)[1:10]
hotel$company_new<-hotel$company
hotel$company_new[hotel$company!="NULL" & hotel$company!="40" & hotel$company!="223" & hotel$company!="67" & hotel$company!="45" & hotel$company!="153" & hotel$company!="174" & hotel$company!="219" & hotel$company!="281" & hotel$company!="154"]<-"OTHER"
table(hotel$company_new)
hotel$company<-hotel$company_new
hotel$company_new<-NULL
hotel$company<-droplevels(hotel$company)


```

Reservation Date
Split into month, day, year
```{r}
library(lubridate)
hotel$res_year <- year(ymd(hotel$reservation_status_date))
hotel$res_month <- month(ymd(hotel$reservation_status_date)) 
hotel$res_day <- day(ymd(hotel$reservation_status_date))
hotel$reservation_status_date<-NULL
hotel$res_year<-as.factor(hotel$res_year)
hotel$res_day<-as.factor(hotel$res_day)
hotel$res_month<-as.factor(hotel$res_month)